A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-10-28 kobakian-cook.Rmd
Changes with Di

The algorithm is named for the *Trigona carbonaria* bee species. Native to northern and eastern Australia this stingless species builds flat layers of hexagonal brood cells, spiralling out from a central point [@PH]. This hive design inspired the use of multiple focal points in the algorithm, individual spirals are constructed by building out from various points on the geographic map base.

The result of the algorithm is a display that emphasises the densely populated geographic regions surrounding the capital cities in Australia.
Each capital city forms it's own collection of connected hexagons. Figure \ref{fig:hexmap} allows the distributions in the capital cities to be viewed at a deeper scale. Sydney is now notable for the considerable amount of red hexagons, not just for the surrounding orange areas around the New South Wales capital visible in the choropleth display in Figure \ref{fig:choro}. Brisbane and Perth also draw attention for the many orange hexagons. Interestingly there is no clear divide between rural and urban SA2 areas as many rural areas, and the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of thyroid cancer diagnosis for females.

@@

There are three key steps performed to create the tessellated hexagon tile map shown in Figure \ref{fig:hexmap}. First, derive the set of centroids from the polygons provided, then create the grid of hexagons locations. These two processes are defined in the blue left column of the flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. There are several filter steps to speed the process of selecting an appropriate hexagon to represent each geographic area. To make tessellated plots with the hexagon allocations, the point locations are converted to hexagon shapes, with the hexagon size used in the grid.

@@

Only two inputs are necessary to begin using the algorithm in the `sugarbag` package; the shape file, and the id variable. The id variable should uniquely identify each geographic unit in the shape file.

The centroids are derived from the shape file. The amount of centroids within the geographic area is used to determine a hexagon size if one is not provided. The derived centroids are a necessary input for an appropriate grid to be constructed, each grid point is located a hexagon size distance from the next closest point in all six directions. The grid will initially cover the entire map space, encompassing all the centroid locations and will extend in all directions to the extent of the buffer distance. This buffer distance can be helpful to account for densely populated coastal areas, allowing the use of the sea for hexagon locations.

The centroid set derived from the shape file is necessary when creating a grid. The hexagon tile map grid requires the creation of the full grid followed by disregarding the points unlikely to be used in the hexagon map.

@@

When allocating representative hexagons, the width parameter can be used to determine the flexibility of positioning. Using the relationship with the nearest reference location, a larger width parameter will increase the amount of available hexagon locations nearer to the centroid of the geographic location. A smaller width will maintain the orientation from the focal point to the centroid when selecting the hexagon location. However, this could mean it is placed further from the geographic centroid location.

@@

The following code creates the hexagon tile map for all the Statistical Areas at Level 2 in Tasmania. These steps can be executed by the main function, `create_hexmap`, or can be run separately for more flexibility.

If a user would like to perform each of the steps of the algorithm themselves, the necessary inputs will change for each function. For example, the `allocate` function can be used directly when a user wishes to use a set of centroids, rather than [@sf] polygons.

The polygon set of Statistical Areas at Level 2 (SA2) [@abs2011] of Tasmania in 2011 is accessed from the `absmapsdata` package. A single column of the data set is used to identify the unique areas. In this case, the unique SA2 names for each SA2 have been used.

The longitude and latitude centre of the capital cities of Australia are used as focal points to allocate each geographic area around the closest capital city. Hobart will be the focal point, as this example uses only the areas in the state of Tasmania.

@@

The buffer distance, hexagon size, hexagon amount to filter and width of angle are parameters that will be determined within `create_hexmap` if they are not provided. They are created as they are needed throughout the following example.

@@

The set of polygons should be provided as an `sf` object, this is a data frame containing a `geometry` column. The `read_shape` function can assist in creating this object for use in `R`.

The centroids can be derived from the set of polygons using the `create_centroids` function:

@@

A grid is created to ensure tessellation between the hexagons that represent the geographic units on a hexagon tile map. The grid of possible hexagon locations is made using the `create_grid` function. It uses the centroids, the hexagon size and the buffer distance.

@@

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by `hex_size`. Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance. A unique hexagon location is created from all intersections of the longitude columns and latitude rows. Figure \ref{fig:filterprocess}a) allows for tessellating hexagons after every second latitude row on the grid is shifted right, by half of the hexagon size.

@@

#### Rolling windows

@@

The closest latitude row and longitude column are found for each centroid location. Then rows and columns of centroids are divided into 20 groups. The amount of rows in each latitude group and the amount of columns in each longitude group are used as the width of rolling windows.

The first rolling window step finds the minimum and maximum centroid values within each of the sliding window groups of longitude columns, and the groups of latitude rows.

The second rolling window step finds the average of the rolling minimum and maximum centroid values, for the longitude columns and latitude rows.

The hexagon grid points are kept only if they fall within the rolling average of the minimum and maximum centroid values after accounting for the buffer distance, for each row and column of the grid. The sparsely populated South-West region of National Park has very few points available compared to the South-East region near the city of Hobart.

@@

The distance is calculated between each centroid in the set, and each of the focal points provided. The distance and angle from the closest focal point to polygon centroid is used in the allocation process as the order for allocation is determined by the distance between the polygon centroid and it's closest focal point.

@@

f2 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas %>% filter(lat < -40.3), fill = "#D3D3D3", colour = "white", size = 0.01) +
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011)), data = fort_allocated, fill = "#353b38", colour = "black", size = 0.01) +
  #Hobart
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "#e9f2ed", size = 3, shape = 4) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, fill = "#ede187", shape = 25, size = 2) + 
  theme_void() + coord_equal() +
  # possible choices to select
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", data = f_grid, size = 1.5) +
  # selected hexagon
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", fill = "#ede187", data = hex, size = 2.5, shape = 21)

@@

A choropleth map display is the current practice spatial visualisation for geographical data. It is possible to enhance the communication of spatial distributions using alternative map displays. Current methods for alternative maps do not always work for Australia due to the large rural geographic space between the densely populated capital cities. The administrative boundaries may also distract from the statistics communicated using colour due to the small size of some areas.

Alternative maps highlight the value of the statistics across the geographic units. These methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each population area is equally important. The solution to this visualisation problem can be using equally sized areas, with neighbourhood boundary connections. The `sugarbag` package for `R` creates tessellated hexagon tile maps by automating an algorithm to design these displays. The Australian application preserves the spatial relationships, emphasising capital cities.

@@

@@

The Dorling cartogram was the most influential alternative map display for the design of the hexagon tile map. The use of a hexagon shifts the focus of the visualisation from the shape of SA2s to the distribution effects of a disease on communities. It is not impeded by the irregular and unusual shapes created by a contiguous cartogram and the immense white space created on the non-contiguous cartogram. Hexagons can emphasise the relationship between neighbouring communities by allowing their sides to touch. The tessellation created by adjoining hexagons for densely populated areas is helpful as it shows that most communities are not islands onto themselves, but are intrinsically connected to those around them. As seen in the Dorling cartogram in figure \ref{fig:tas_displays}c), the tessellation that allows boundary connections was not imposed for rural communities, where centroid of the geographic unit was located far from others, a feature that allows distant but dense populations of the coastal towns of southern Tasmania to be recognised.

<!-- future -->

This algorithm has only been tested using single countries, and does not consider definite borders of countries. While the buffer allows extension beyond the furthest centroids, there is no mechanism to protect the borders and ensure centroids are placed within the geographic borders of a country.
It would be also possible to take a logarithmic function in the filter steps for selection of a hexagon grid point. A direct angle is currently used, but this may narrowly miss appropriate hexagons, a logarithmic function may help to choose a closer hexagon to the original centroid location, before increasing the width of the angle used to filter the hexagons.
This algorithm is an effective start to creating hexagon tile maps for many geographic units.

### Installation

>>>>

The algorithm is implemented in the `sugarbag` package for R, this is the common name for the *Trigona carbonaria* bee species. Native to northern and eastern Australia this stingless species builds flat layers of hexagonal brood cells, spiralling out from a central point [@PH]. This hive design inspired the use of multiple focal points in the algorithm, individual spirals are constructed by building out from various points on the geographic map base.

The result of the algorithm on rates of thyroid cancer for females in each of the SA2 areas of Australia is shown in Figure \ref{fig:hexmap}.
The difference to the choropleth map in Figure \ref{fig:choro} are clear. Now the thyroid cancer incidence in the densely populated areas in Sydney, Brisbane and Perth are visible. Interestingly there is no clear divide between rural and urban SA2 areas as many rural areas, and the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of thyroid cancer diagnosis for females. The hexagon tile map display provides a more accurate representation of the spatial distribution of thyroid cancer incidence across Australia.

@@

There are several key steps needed to create the hexagon tile map as described in the flow chart in Figure \ref{fig:sugarbag_flow}. First, derive the set of centroids from the polygons provided, then create the grid of hexagon locations. These two processes are defined in the blue left column of the flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. There are several filter steps to speed the process of selecting an appropriate hexagon to represent each geographic unit. To make tessellated plots with the hexagon allocations, the point locations are converted into hexagon shapes.

@@

Only two inputs are necessary to begin the algorithm; the shape file, and the id variable. The id variable should uniquely identify each geographic unit in the shape file.

The centroids are derived from the shape file. The amount of centroids within the geographic map is used to determine an appropriate hexagon size if one is not provided. The derived centroids are a necessary input for an appropriate grid to be constructed. Each grid point is located a hexagon size distance from the next closest point in all six directions. The grid will initially cover the entire map space, encompassing all the centroid locations and will extend in all directions to the extent of the buffer distance. This buffer distance can be helpful to account for densely populated coastal areas, allowing the use of the sea for hexagon locations.

@@

When allocating representative hexagons, the width parameter can be used to determine the flexibility of positioning. Using the relationship with the nearest focal point, a larger width parameter will increase the amount of available hexagons nearer to the original centroid of the geographic unit. A smaller width will maintain the orientation from the focal point to the centroid when selecting the hexagon location. However, this could mean it is placed further away.

@@

## Algorithm steps

The following steps create the hexagon tile map for all the Statistical Areas at Level 2 in Tasmania. These steps can be executed by the main function, `create_hexmap`, or can be run separately for more flexibility.

If a user would like to perform each of the steps of the algorithm themselves, the necessary inputs will change for each function. For example, the `allocate` function can be used directly when a user wishes to use a set of centroids, rather than polygons.

The polygon set of Tasmania in 2011 is accessed from the `absmapsdata` package. A single column of the data set is used to identify the unique areas. In this case, the unique SA2 names for each SA2 have been used.

The Australian capital cities are used as focal points to allocate each geographic area around the closest capital city. Hobart will be the focal point for this example because only the state of Tasmania is being processed.

@@

The buffer distance, hexagon size, hexagon amount to filter and width of angle are parameters that will be determined within `create_hexmap()`, if they are not provided. They are created as they are needed throughout the following example.

The results from various steps in the process are illustrated in Figure 5.

@@

The set of polygons should be provided as an `sf` object, this is a data frame containing a `geometry` column. The `read_shape()` function can assist in creating this object for use in `R`.

The centroids can be derived from an `sf` object using the `create_centroids()` function:

@@

A grid is created to ensure tessellation between the hexagons that represent the geographic units on a hexagon tile map. The grid of possible hexagon locations is made using the `create_grid()` function. It uses the centroids, the hexagon size and the buffer distance.

@@

A set of longitude columns, and latitude rows are created to define the locations of the hexagons. The distance between each row and column is the size specified by `hex_size`. Equally spaced columns are created from the minimum longitude minus the buffer distance, up to the maximum longitude plus the buffer distance. Similarly, the rows are created from the latitude values and the buffer distance. A unique hexagon location is created from all intersections of the longitude columns and latitude rows. Figure \ref{fig:filterprocess}a) shows the hexagon grid after every second latitude row on the grid is shifted right, by half of the hexagon size.

@@

@@

The closest latitude row and longitude column are found for each centroid location. Then rows and columns of centroids are divided into 20 groups. The amount of rows in each latitude group and the amount of columns in each longitude group are used as the width of rolling windows. The first rolling window step finds the minimum and maximum centroid values within each of the sliding window groups of longitude columns, and the groups of latitude rows. The second rolling window step finds the average of the rolling minimum and maximum centroid values, for the longitude columns and latitude rows.

The hexagon grid points are kept only if they fall within the rolling average of the minimum and maximum centroid values after accounting for the buffer distance, for each row and column of the grid. Figure 5b displays remaining hexagon grid points after applying the buffer step. The sparsely populated South-West region of National Park has very few points available compared to the South-East region near the city of Hobart.

@@

The distance is calculated between each centroid in the set, and each of the focal points provided. The distance and angle from the closest focal point to polygon centroid is used in the allocation process as the order for allocation is determined by the distance between the polygon centroid and it's closest focal point. For this example, this distance is only calculated for the capital city of Hobart.

@@

f2 <- ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011, polygon)), data = fort_tas %>% filter(lat < -40.3), fill = "#D3D3D3", colour = "white", size = 0.01) +
  geom_polygon(aes(x = long, y = lat, group = interaction(sa2_name_2011)), data = fort_allocated, fill = "#353b38", colour = "black", size = 0.01) +
  #Hobart
  geom_point(aes(x = focal_longitude, y = focal_latitude), data = f_centroid, colour = "#e9f2ed", size = 3, shape = 4) + 
  geom_point(aes(x=longitude, y = latitude), data = f_centroid, fill = "#ede187", shape = 25, size = 2) + 
 # theme_void() + coord_equal() +
  # possible choices to select
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", data = f_grid, size = 1.5) +
  # selected hexagon
  geom_point(aes(x = hex_long, y = hex_lat), colour = "#09661f", fill = "#ede187", data = hex, size = 2.5, shape = 21)

@@

<!-- Summarise paper, usability -->
A choropleth map display is the current common practice for spatial data visualisation. As the choropleth map display relies on the geographic shapes, it does not effectively present the density of the population in small geographic areas. The common solution to enhance the communication of spatial distributions is to incorporate the population values as a cartogram or another alternative map display. Many alternative map displays do not work well for Australia due to the large rural geographic space between the densely populated capital cities. The administrative boundaries may also distract from the statistics communicated using colour due to the small size of some areas.
This paper provides an algorithm for Australia called the hexagon tile map.

<!-- implementation -->
Alternative maps highlight the value of the statistics across the geographic units. These methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing each administrative area. This acknowledges that the amount of residents can be different but recognises that each population area is equally important. The solution to this visualisation problem can be using equally sized areas, with neighbourhood boundary connections. The `sugarbag` package for `R` creates tessellated hexagon tile maps by automating an algorithm to design these displays. The Australian application preserves the spatial relationships, emphasising capital cities.

@@

This algorithm should  be general enough to apply to other countries and problems

@@

<!-- The Dorling cartogram was the most influential alternative map display for the design of the hexagon tile map. The use of a hexagon shifts the focus of the visualisation from the shape of SA2s to the distribution effects of a disease on communities. It is not impeded by the irregular and unusual shapes created by a contiguous cartogram and the immense white space created on the non-contiguous cartogram. Hexagons can emphasise the relationship between neighbouring communities by allowing their sides to touch. The tessellation created by adjoining hexagons for densely populated areas is helpful as it shows that most communities are not islands onto themselves, but are intrinsically connected to those around them. As seen in the Dorling cartogram in figure \ref{fig:tas_displays}c), the tessellation that allows boundary connections was not imposed for rural communities, where centroid of the geographic unit was located far from others, a feature that allows distant but dense populations of the coastal towns of southern Tasmania to be recognised. -->

<!-- , limitiations, where it might not work -->

This algorithm has only been tested using single countries, and does not consider definite borders of countries. While the buffer allows extension beyond the furthest centroids, there is no mechanism to protect the borders and ensure centroids are placed within the geographic borders of a country.

<!-- future -->
It would be also possible to take a logarithmic function in the filter steps for selection of a hexagon grid point. A direct angle is currently used, but this may narrowly miss appropriate hexagons, a logarithmic function may help to choose a closer hexagon to the original centroid location, before increasing the width of the angle used to filter the hexagons.
This algorithm is an effective start to creating hexagon tile maps for many geographic units.

### 2020-10-28 kobakian-cook.tex
Changes with Di

The algorithm is named for the \emph{Trigona carbonaria} bee species.
Native to northern and eastern Australia this stingless species builds
flat layers of hexagonal brood cells, spiralling out from a central
point \citep{PH}. This hive design inspired the use of multiple focal
points in the algorithm, individual spirals are constructed by building
out from various points on the geographic map base.

The result of the algorithm is a display that emphasises the densely
populated geographic regions surrounding the capital cities in
Australia. Each capital city forms it's own collection of connected
hexagons. Figure \ref{fig:hexmap} allows the distributions in the
capital cities to be viewed at a deeper scale. Sydney is now notable for
the considerable amount of red hexagons, not just for the surrounding
orange areas around the New South Wales capital visible in the
choropleth display in Figure \ref{fig:choro}. Brisbane and Perth also
draw attention for the many orange hexagons. Interestingly there is no
clear divide between rural and urban SA2 areas as many rural areas, and
the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of
thyroid cancer diagnosis for females.

@@

Only two inputs are necessary to begin using the algorithm in the
\texttt{sugarbag} package; the shape file, and the id variable. The id
variable should uniquely identify each geographic unit in the shape
file.

The centroids are derived from the shape file. The amount of centroids
within the geographic area is used to determine a hexagon size if one is
not provided. The derived centroids are a necessary input for an
appropriate grid to be constructed, each grid point is located a hexagon
size distance from the next closest point in all six directions. The
grid will initially cover the entire map space, encompassing all the
centroid locations and will extend in all directions to the extent of
the buffer distance. This buffer distance can be helpful to account for
densely populated coastal areas, allowing the use of the sea for hexagon
locations.

The centroid set derived from the shape file is necessary when creating
a grid. The hexagon tile map grid requires the creation of the full grid
followed by disregarding the points unlikely to be used in the hexagon
map.

@@

The following code creates the hexagon tile map for all the Statistical
Areas at Level 2 in Tasmania. These steps can be executed by the main
function, \texttt{create\_hexmap}, or can be run separately for more
flexibility.

If a user would like to perform each of the steps of the algorithm
themselves, the necessary inputs will change for each function. For
example, the \texttt{allocate} function can be used directly when a user
wishes to use a set of centroids, rather than \citep{sf} polygons.

The polygon set of Statistical Areas at Level 2 (SA2) \citep{abs2011} of
Tasmania in 2011 is accessed from the \texttt{absmapsdata} package. A
single column of the data set is used to identify the unique areas. In
this case, the unique SA2 names for each SA2 have been used.

The longitude and latitude centre of the capital cities of Australia are
used as focal points to allocate each geographic area around the closest
capital city. Hobart will be the focal point, as this example uses only
the areas in the state of Tasmania.

The buffer distance, hexagon size, hexagon amount to filter and width of
angle are parameters that will be determined within
\texttt{create\_hexmap} if they are not provided. They are created as
they are needed throughout the following example.

@@

The set of polygons should be provided as an \texttt{sf} object, this is
a data frame containing a \texttt{geometry} column. The
\texttt{read\_shape} function can assist in creating this object for use
in \texttt{R}.

The centroids can be derived from the set of polygons using the
\texttt{create\_centroids} function:

@@

A grid is created to ensure tessellation between the hexagons that
represent the geographic units on a hexagon tile map. The grid of
possible hexagon locations is made using the \texttt{create\_grid}
function. It uses the centroids, the hexagon size and the buffer
distance.

@@

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
the size specified by \texttt{hex\_size}. Equally spaced columns are
created from the minimum longitude minus the buffer distance, up to the
maximum longitude plus the buffer distance. Similarly, the rows are
created from the latitude values and the buffer distance. A unique
hexagon location is created from all intersections of the longitude
columns and latitude rows. Figure \ref{fig:filterprocess}a) allows for
tessellating hexagons after every second latitude row on the grid is
shifted right, by half of the hexagon size.

\hypertarget{rolling-windows}{%
\paragraph{Rolling windows}\label{rolling-windows}}

@@

A choropleth map display is the current practice spatial visualisation
for geographical data. It is possible to enhance the communication of
spatial distributions using alternative map displays. Current methods
for alternative maps do not always work for Australia due to the large
rural geographic space between the densely populated capital cities. The
administrative boundaries may also distract from the statistics
communicated using colour due to the small size of some areas.

@@

The Dorling cartogram was the most influential alternative map display
for the design of the hexagon tile map. The use of a hexagon shifts the
focus of the visualisation from the shape of SA2s to the distribution
effects of a disease on communities. It is not impeded by the irregular
and unusual shapes created by a contiguous cartogram and the immense
white space created on the non-contiguous cartogram. Hexagons can
emphasise the relationship between neighbouring communities by allowing
their sides to touch. The tessellation created by adjoining hexagons for
densely populated areas is helpful as it shows that most communities are
not islands onto themselves, but are intrinsically connected to those
around them. As seen in the Dorling cartogram in figure
\ref{fig:tas_displays}c), the tessellation that allows boundary
connections was not imposed for rural communities, where centroid of the
geographic unit was located far from others, a feature that allows
distant but dense populations of the coastal towns of southern Tasmania
to be recognised.

This algorithm has only been tested using single countries, and does not
consider definite borders of countries. While the buffer allows
extension beyond the furthest centroids, there is no mechanism to
protect the borders and ensure centroids are placed within the
geographic borders of a country. It would be also possible to take a
logarithmic function in the filter steps for selection of a hexagon grid
point. A direct angle is currently used, but this may narrowly miss
appropriate hexagons, a logarithmic function may help to choose a closer
hexagon to the original centroid location, before increasing the width
of the angle used to filter the hexagons. This algorithm is an effective
start to creating hexagon tile maps for many geographic units.

\hypertarget{installation}{%
\subsubsection{Installation}\label{installation}}

>>>>

The algorithm is implemented in the \texttt{sugarbag} package for R,
this is the common name for the \emph{Trigona carbonaria} bee species.
Native to northern and eastern Australia this stingless species builds
flat layers of hexagonal brood cells, spiralling out from a central
point \citep{PH}. This hive design inspired the use of multiple focal
points in the algorithm, individual spirals are constructed by building
out from various points on the geographic map base.

The result of the algorithm on rates of thyroid cancer for females in
each of the SA2 areas of Australia is shown in Figure \ref{fig:hexmap}.
The difference to the choropleth map in Figure \ref{fig:choro} are
clear. Now the thyroid cancer incidence in the densely populated areas
in Sydney, Brisbane and Perth are visible. Interestingly there is no
clear divide between rural and urban SA2 areas as many rural areas, and
the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of
thyroid cancer diagnosis for females. The hexagon tile map display
provides a more accurate representation of the spatial distribution of
thyroid cancer incidence across Australia.

@@

Only two inputs are necessary to begin the algorithm; the shape file,
and the id variable. The id variable should uniquely identify each
geographic unit in the shape file.

The centroids are derived from the shape file. The amount of centroids
within the geographic map is used to determine an appropriate hexagon
size if one is not provided. The derived centroids are a necessary input
for an appropriate grid to be constructed. Each grid point is located a
hexagon size distance from the next closest point in all six directions.
The grid will initially cover the entire map space, encompassing all the
centroid locations and will extend in all directions to the extent of
the buffer distance. This buffer distance can be helpful to account for
densely populated coastal areas, allowing the use of the sea for hexagon
locations.

@@

\hypertarget{algorithm-steps}{%
\subsection{Algorithm steps}\label{algorithm-steps}}

The following steps create the hexagon tile map for all the Statistical
Areas at Level 2 in Tasmania. These steps can be executed by the main
function, \texttt{create\_hexmap}, or can be run separately for more
flexibility.

If a user would like to perform each of the steps of the algorithm
themselves, the necessary inputs will change for each function. For
example, the \texttt{allocate} function can be used directly when a user
wishes to use a set of centroids, rather than polygons.

The polygon set of Tasmania in 2011 is accessed from the
\texttt{absmapsdata} package. A single column of the data set is used to
identify the unique areas. In this case, the unique SA2 names for each
SA2 have been used.

The Australian capital cities are used as focal points to allocate each
geographic area around the closest capital city. Hobart will be the
focal point for this example because only the state of Tasmania is being
processed.

The buffer distance, hexagon size, hexagon amount to filter and width of
angle are parameters that will be determined within
\texttt{create\_hexmap()}, if they are not provided. They are created as
they are needed throughout the following example.

The results from various steps in the process are illustrated in Figure
5.

@@

The set of polygons should be provided as an \texttt{sf} object, this is
a data frame containing a \texttt{geometry} column. The
\texttt{read\_shape()} function can assist in creating this object for
use in \texttt{R}.

The centroids can be derived from an \texttt{sf} object using the
\texttt{create\_centroids()} function:

@@

A grid is created to ensure tessellation between the hexagons that
represent the geographic units on a hexagon tile map. The grid of
possible hexagon locations is made using the \texttt{create\_grid()}
function. It uses the centroids, the hexagon size and the buffer
distance.

@@

A set of longitude columns, and latitude rows are created to define the
locations of the hexagons. The distance between each row and column is
the size specified by \texttt{hex\_size}. Equally spaced columns are
created from the minimum longitude minus the buffer distance, up to the
maximum longitude plus the buffer distance. Similarly, the rows are
created from the latitude values and the buffer distance. A unique
hexagon location is created from all intersections of the longitude
columns and latitude rows. Figure \ref{fig:filterprocess}a) shows the
hexagon grid after every second latitude row on the grid is shifted
right, by half of the hexagon size.

@@

A choropleth map display is the current common practice for spatial data
visualisation. As the choropleth map display relies on the geographic
shapes, it does not effectively present the density of the population in
small geographic areas. The common solution to enhance the communication
of spatial distributions is to incorporate the population values as a
cartogram or another alternative map display. Many alternative map
displays do not work well for Australia due to the large rural
geographic space between the densely populated capital cities. The
administrative boundaries may also distract from the statistics
communicated using colour due to the small size of some areas. This
paper provides an algorithm for Australia called the hexagon tile map.

@@

This algorithm should be general enough to apply to other countries and
problems

This algorithm has only been tested using single countries, and does not
consider definite borders of countries. While the buffer allows
extension beyond the furthest centroids, there is no mechanism to
protect the borders and ensure centroids are placed within the
geographic borders of a country.

It would be also possible to take a logarithmic function in the filter
steps for selection of a hexagon grid point. A direct angle is currently
used, but this may narrowly miss appropriate hexagons, a logarithmic
function may help to choose a closer hexagon to the original centroid
location, before increasing the width of the angle used to filter the
hexagons. This algorithm is an effective start to creating hexagon tile
maps for many geographic units.
